---
title: "p8105_hw6_RYW2109"
author: "Rita Wang"
date: "2024-12-02"
output: github_document
---
```{r default_setting, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
library(tidyverse)
library(ggplot2)
library(purrr)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

```{r q1_data, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
data_file = "./data"

if (!dir.exists(data_file)) {
  dir.create(data_file)
}
weather_df = 
  rnoaa::meteo_pull_monitors(
      # install.packages("rnoaa") --> "package ‘rnoaa’ is not available for this version of R"
      # 1. install.packages("remotes")
      # 2. remotes::install_github("ropensci/rnoaa")
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything()) 

# Save your data frame to the new directory
file_path = file.path(data_file, "weather_data.csv")
write.csv(weather_df, file_path, row.names = FALSE)

weather_df
```

```{r q1_a, echo = FALSE, message = FALSE, warning = FALSE}
# Bootstrap function
bootstrap_results = function(data, n_bootstrap = 5000) {
  boot_results = replicate(n_bootstrap, {
    sample_data = sample_n(data, size = nrow(data), replace = TRUE)
    model = lm(tmax ~ tmin, data = sample_data)
      # Fit linear regression model
    r_squared = broom::glance(model)$r.squared # R^2
    # log(beta0 * beta1)
    coeffs = broom::tidy(model)
    beta0 = coeffs$estimate[1]
    beta1 = coeffs$estimate[2]
    log_beta_product = log(abs(beta0 * beta1))
    c(r_squared = r_squared, log_beta_product = log_beta_product)
  })
  
  # Convert results into a data frame
  boot_results_df = as.data.frame(t(boot_results))
    colnames(boot_results_df) = c("r_squared", "log_beta_product")
    return(boot_results_df)
}

# Bootstrap with 5000 samples
bootstrap_df = bootstrap_results(weather_df, n_bootstrap = 5000)

str(bootstrap_df)

# r_squared plot
ggplot(bootstrap_df, aes(x = r_squared)) +
  geom_histogram(bins = 50, fill = "purple", color = "red") +
  labs(title = "Distribution of R-squared estimates from bootstrapping",
       x = "R-squared", y = "Frequency")

# log_beta_product
ggplot(bootstrap_df, aes(x = log_beta_product)) +
  geom_histogram(bins = 50, fill = "orange", color = "red") +
  labs(title = "Distribution of log(Beta0 * Beta1) estimates from bootstrapping",
       x = "log(Beta0 * Beta1)", y = "Frequency")

```
The R-squared distribution has a normal distribution, which also represents the percentage of variance in tmax that is explained by tmin. The log(Beta0 * Beta1) distribution also has a relative normal distribution. Beta0 represents the intercept of the regression model with Beta1 representing the slope; the product of these two values in the distribution provides an idea of the uncertainty of the interaction between the intercept and slope.

```{r q1_b, echo = FALSE, message = FALSE, warning = FALSE}
ci_r_squared = quantile(bootstrap_results$r_squared, conf_interval = c(0.025, 0.975))
ci_log_beta = quantile(bootstrap_results$log_beta, conf_interval = c(0.025, 0.975))

# Print the confidence intervals
print("95% CI for R-squared:")
print(r_squared_ci)

print("95% CI for log(Beta0 * Beta1):")
print(log_beta_product_ci)
```

```{r q2_data, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
  # importing data
url = "https://raw.githubusercontent.com/washingtonpost/data-homicides/refs/heads/master/homicide-data.csv"

file_path = file.path(data_file, "homicide-data.csv")
download.file(url, destfile = file_path, method = "libcurl")

data_original = read.csv(file_path) %>% 
  janitor::clean_names()
```

```{r q2_a, echo = FALSE, message = FALSE, warning = FALSE}
data_df1 = data_original %>% 
  mutate(
    city_state = paste(city, state, sep = ", ")
  )

data_df2 = data_df1 %>%
  group_by(city_state) %>%
  summarise(
    total_homicides = n(),
    unsolved_homicides = sum(disposition %in% c("Closed without arrest", "Open/No arrest")),
    .groups = "drop"
  )
```
