---
title: "p8105_hw6_RYW2109"
author: "Rita Wang"
date: "2024-12-02"
output: github_document
---
```{r default_setting, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
library(tidyverse)
library(ggplot2)
library(purrr)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

```{r q1_data, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
data_file = "./data"

if (!dir.exists(data_file)) {
  dir.create(data_file)
}
weather_df = 
  rnoaa::meteo_pull_monitors(
      # install.packages("rnoaa") --> "package ‘rnoaa’ is not available for this version of R"
      # 1. install.packages("remotes")
      # 2. remotes::install_github("ropensci/rnoaa")
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything()) 

# Save your data frame to the new directory
file_path = file.path(data_file, "weather_data.csv")
write.csv(weather_df, file_path, row.names = FALSE)

weather_df
```

```{r q1_a, echo = FALSE, message = FALSE, warning = FALSE}
# Bootstrap function
bootstrap_results = function(data, n_bootstrap = 5000) {
  boot_results = replicate(n_bootstrap, {
    sample_data = sample_n(data, size = nrow(data), replace = TRUE)
    model = lm(tmax ~ tmin, data = sample_data)
      # Fit linear regression model
    r_squared = broom::glance(model)$r.squared # R^2
    # log(beta0 * beta1)
    coeffs = broom::tidy(model)
    beta0 = coeffs$estimate[1]
    beta1 = coeffs$estimate[2]
    log_beta_product = log(abs(beta0 * beta1))
    c(r_squared = r_squared, log_beta_product = log_beta_product)
  })
  
  # Convert results into a data frame
  boot_results_df = as.data.frame(t(boot_results))
    colnames(boot_results_df) = c("r_squared", "log_beta_product")
    return(boot_results_df)
}

# Bootstrap with 5000 samples
bootstrap_df = bootstrap_results(weather_df, n_bootstrap = 5000)

str(bootstrap_df)

# r_squared plot
ggplot(bootstrap_df, aes(x = r_squared)) +
  geom_histogram(bins = 50, fill = "purple", color = "red") +
  labs(title = "Distribution of R-squared estimates from bootstrapping",
       x = "R-squared", y = "Frequency")

# log_beta_product
ggplot(bootstrap_df, aes(x = log_beta_product)) +
  geom_histogram(bins = 50, fill = "orange", color = "red") +
  labs(title = "Distribution of log(Beta0 * Beta1) estimates from bootstrapping",
       x = "log(Beta0 * Beta1)", y = "Frequency")

```
The R-squared distribution has a normal distribution, which also represents the percentage of variance in tmax that is explained by tmin. The log(Beta0 * Beta1) distribution also has a relative normal distribution. Beta0 represents the intercept of the regression model with Beta1 representing the slope; the product of these two values in the distribution provides an idea of the uncertainty of the interaction between the intercept and slope.

```{r q1_b, echo = FALSE, message = FALSE, warning = FALSE}
ci_r_squared = quantile(bootstrap_df$r_squared, probs = c(0.025, 0.975))
ci_log_beta = quantile(bootstrap_df$log_beta, probs = c(0.025, 0.975))

# Print the confidence intervals
print("95% CI for R-squared:")
print(ci_r_squared)

print("95% CI for log(Beta0 * Beta1):")
print(ci_log_beta)
```

```{r q2_data, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
  # importing data
url = "https://raw.githubusercontent.com/washingtonpost/data-homicides/refs/heads/master/homicide-data.csv"

file_path = file.path(data_file, "homicide-data.csv")
download.file(url, destfile = file_path, method = "libcurl")

data_original = read.csv(file_path) %>% 
  janitor::clean_names()
```

```{r q2_a, echo = FALSE, message = FALSE, warning = FALSE}
data_df1 = data_original %>%
  filter(!(city %in% c("Dallas", "Phoenix", "Kansas City", "Tulsa") &
             state %in% c("TX", "AZ", "MO", "AL"))) %>%
  filter(victim_race %in% c("White", "Black")) %>% 
  mutate(
    homicide_solved = ifelse(grepl("Closed", disposition, ignore.case = TRUE), 1, 0),
      # 1 --> closed --> solved
      # 0 --> open --> not solved
    victim_age = as.numeric(victim_age)
  )

cleaned = data_df1 %>% 
  mutate(
    city_state = paste(city, state, sep = ", "),
  )
```

```{r q2_b, echo = FALSE, message = FALSE, warning = FALSE}
baltimore_data = cleaned %>%
  filter(city_state == "Baltimore, MD")

# Logistic regression model using glm()
log_reg_model = glm(homicide_solved ~ victim_age + victim_sex + victim_race,
                    data = baltimore_data,
                    family = "binomial")

# Tidy summary of the model
model_summary = broom::tidy(log_reg_model, conf.int = TRUE)
print(model_summary)

# Odds ratio and 95% CI for sex
sex_odds_ratio = model_summary %>%
  filter(term == "victim_sexMale") %>% 
  select(estimate, conf.low, conf.high)
print(sex_odds_ratio)
```

```{r q2_c, echo = FALSE, message = FALSE, warning = FALSE}
# Logistic regression model using glm()
get_odds_ratio = function(city_data) {
  log_reg_model = glm(homicide_solved ~ victim_age + victim_sex + victim_race,
                    data = city_data,
                    family = "binomial")

# Tidy summary of the model
model_summary = broom::tidy(log_reg_model, conf.int = TRUE)
print(model_summary)

# Odds ratio and 95% CI for sex
sex_odds_ratio = model_summary %>%
  filter(term == "victim_sexMale") %>%
  mutate(
    odds_ratio = exp(estimate),
    conf.low = exp(conf.low),
    conf.high = exp(conf.high)
    ) %>%
  select(odds_ratio, conf.low, conf.high)
}

# For each city
city_results = cleaned %>%
  group_by(city_state) %>%
  nest() %>%  # List-column city
  mutate(
    odds_ratios = map(data, get_odds_ratio)
  ) %>%
  select(city_state, odds_ratios) %>%
  unnest(odds_ratios)
```

```{r q2_d, echo = FALSE, message = FALSE, warning = FALSE}
ggplot(city_results, aes(x = odds_ratio, y = reorder(city_state, odds_ratio))) +
  geom_point(aes(color = odds_ratio), size = 4) +  # Plot OR as points
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  labs(
    title = "Estimated Odds Ratios (ORs) for Solving Homicides: Male vs Female Victims",
    x = "Odds Ratio",
    y = "City",
    caption = "95% Confidence Intervals shown as error bars"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),
    plot.title = element_text(hjust = 0.5),
    plot.caption = element_text(hjust = 0.5)
  ) +
  scale_color_gradient(low = "red", high = "blue")
```

```{r q3_data, default_setting, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
# importing birthweight
birth_original = read_csv(file = "./Data/birthweight.csv") %>% 
  janitor::clean_names()